---
title: "T6SS_post_analysis"
author: "Hongxi"
date: "`r Sys.Date()`"
output: html_document
---
#data preparation
load required packages

```{r preparation, message=FALSE}
library(tidyr)
library(readxl)
library(gggenes)
library(ggplot2)
library(ggfittext)
library(dplyr)
```

load and wash the T6SS blast data

```{r data loading, echo=FALSE, warning=TRUE}
rm(list = ls())

T6SS_blast_raw <- read.csv("../../Results/T6SS_identification/T6SS_blast_result.csv",header = T, sep = " ")
metadata <- read.csv("../../data/metadata/datasets_used_in_analysis.csv", sep = " ")
available_metadata_list <- metadata["Assembly_Accession"]
locus_info <- read.csv("../../data/processed_data/CJ_RefSeq/all_locus_info.csv")

T6SS_blast_raw$Subject_accession_version <- sub("\\..*", "", T6SS_blast_raw$Subject_accession_version)
T6SS_blast_raw <- merge(T6SS_blast_raw, locus_info, by.x = "Subject_accession_version", by.y = "locus_accession", all.x = T)
T6SS_blast_raw <- merge(T6SS_blast_raw,available_metadata_list, by.x = "accession", by.y = "Assembly_Accession")

```


#determine the T6SS positive results

```{r}
  # calculate the hits' coverage proportion in query protein 
    T6SS_blast_raw$hit_coverage <- T6SS_blast_raw$Alignment_length / T6SS_blast_raw$Query_sequence_length
  
  # select the qualified results from blast results
    T6SS_blast_available <- subset(T6SS_blast_raw, 
                           subset = Percentage_of_identical_matches >=50 & hit_coverage >= 0.5,
                           select = c("Query_accession_version","accession")) 
  
  # calculate the frequency of hits for each protein and clean the data
    T6SS_blast_available <- T6SS_blast_available %>% group_by(Query_accession_version, accession) %>% mutate(frequency = n())
    T6SS_blast_available <- distinct(T6SS_blast_available)
  
  # change the data format from long-data to wide-data
    T6SS_blast_available_wide <- spread(T6SS_blast_available, key = Query_accession_version, value = frequency, fill = 0)
  
  # calculate how many proteins are present in each genome
    T6SS_blast_available_wide$target_score <- rowSums(T6SS_blast_available_wide != 0) -1
  
  # select the presence of proteins that less than 11 in 13 
    T6SS_positive_result <- subset(T6SS_blast_available_wide,
                                 target_score >= 11)
  
  # generate a simpler table of the result
    T6SS_positive_simple <- subset(T6SS_positive_result,
                                 select = c("accession","target_score"))
  
  # combine the T6SS positive result with metadata
    T6SS_with_metadata <- merge(T6SS_positive_simple, metadata, by.x = "accession", by.y = "Assembly_Accession")

```

## output T6SS positive results into Result folder

```{r}
    write.table(T6SS_positive_result, file = "../../Results/T6SS_identification/T6SS_positive_protein_detail.csv" )
    write.table(T6SS_with_metadata, file = "../../Results/T6SS_identification/T6SS_positive_metadata.csv" )
    writeLines(T6SS_positive_simple$accession, "../../Results/T6SS_identification/T6SS_positive_list.txt")
    write.table(T6SS_blast_raw, file = "../../Results/T6SS_identification/T6SS_blast_result_post_analysised.csv", sep = ",", row.names = F)

```

### View the T6SS positive results (optional)

```{r}
View(T6SS_positive_result)
View(T6SS_with_metadata)
View(T6SS_positive_simple)
```


#determine T6SS positive ranges in genome
note the threshold to classify different T6SS locus is manually set as 20000 base (20 kb) 

```{r}
# select qualified results
range_T6SS <- subset(T6SS_blast_raw, 
                     subset = Percentage_of_identical_matches >=50 & hit_coverage >= 0.5) 

# calculate start and end point in genome
range_T6SS$start_in_genome <- range_T6SS$Start_of_alignment_in_subject + range_T6SS$start
range_T6SS$end_in_genome <- range_T6SS$End_of_alignment_in_subject + range_T6SS$start

# select T6SS positive results
range_T6SS_positive <- merge(range_T6SS,T6SS_positive_simple, by = "accession", all.y = T)

# determine the direction for each genes for graph generation 
range_T6SS_positive <- subset(range_T6SS_positive, select = c("Query_accession_version","accession","start_in_genome","end_in_genome"))
range_T6SS_positive$orientation <- ifelse(range_T6SS_positive$end_in_genome < range_T6SS_positive$start_in_genome, "reverse", "forward")
range_T6SS_positive$type <- "T6SS"

# rename col name
colnames(range_T6SS_positive) <- c("sequence_code","accession","start","end","orientation","type")

#build a function to make sure start point < end point
swap_start_end <- function(data) {
  # get the rows that need to be swiped 
  rows_to_swap <- data$orientation == "reverse"
  
  # swap the start and end value 
  data[rows_to_swap, c("start", "end")] <- data[rows_to_swap, c("end", "start")]
  
  # return the swiped data
  return(data)
}
# apply the function
range_T6SS_positive <- swap_start_end(range_T6SS_positive)
# split the data
range_T6SS_list <- split(range_T6SS_positive,range_T6SS_positive$accession)

# group each genes by the criteria:
# any two adjacent genes with a start position difference greater than 20,000 bases are considered to be in different groups
# accessions have more than 2 groups will be removed from the analysis
threshold <- 20000
group_the_ranges <- function(df,threshold) {
  accession <- unique(df$accession)
  df <- arrange(df,start)
  df$group <- cumsum(c(1, diff(df$start) > threshold))
    if (length(unique(df$group)) > 3) {
    cat(paste0(accession, " has more than 3 sequences, dropping.\n"))
    return(NULL)
  } else {
    return(df)
  }
}
# apply the function 
range_T6SS_list <- lapply(range_T6SS_list, group_the_ranges, threshold = threshold)
range_T6SS_list <- range_T6SS_list[!sapply(range_T6SS_list, is.null)]

```
now the element `range_T6SS_list` contains all the T6SS gene's information of start point, end point, direction, molecule type and which groups they are belonging to

## generate every T6SS-positive plots into a list Results folder as .png file

```{r}
# set the function for plots generation 
draw_T6SS <- function(df) {
  T6SS_plot <- ggplot(df, aes(xmin = start, xmax = end, y = orientation, fill = sequence_code, label = sequence_code, forward = orientation == "forward")) +
    geom_gene_arrow() +
    geom_gene_label(align = "left") +
    facet_grid(cols = vars(group), scales = "free", space = "fixed") +
    scale_fill_brewer(palette = "Set3",name = "sequence_code") +
    theme_genes()
  return(T6SS_plot)
}

# apply the function
range_T6SS_plots <- lapply(range_T6SS_list,draw_T6SS)

```

### check one of the plot to view (optional)

```{r warning=FALSE}
print(range_T6SS_plots[[39]])
```


### output the plots into Result folder

```{r}
# output the plots
for (i in seq_along(range_T6SS_plots)) {
  
  plot_name <- names(range_T6SS_plots)[i]
  file_name <- paste0("../../Results/T6SS_identification/T6SS_plots/",plot_name,".png")
  
  ggsave(file_name, plot = range_T6SS_plots[[i]],width = 12, height = 4, device = "png")
}

```
now the T6SS proteins' architecture in positive genomes are generated and storied in target folder 


#determine the larger range for each T6SS gene sets
  this code use a combined function to generate the start and end point for each T6SS gene group.
```{r}
find_all_ranges <- function(data) {
  
  accession_gene_list <- split(data,data$group)
  
  find_the_ranges <- function(df) {
    seq_ranges <- data.frame(accession = unique(df$accession), start = min(df$start), end = max(df$end), type = unique(df$type) , group = unique(df$group))
    return(seq_ranges)
  }
  
  accession_range_list <- lapply(accession_gene_list, find_the_ranges)
  accession_range_list <- do.call(rbind,accession_range_list)
  
  return(accession_range_list)
}

whole_ranges_location <- lapply(range_T6SS_list,find_all_ranges)
whole_ranges_location <- do.call(rbind,whole_ranges_location)
rownames(whole_ranges_location) <- NULL


```

### check the large T6SS range dataframe (optional)

```{r}
print(head(whole_ranges_location, n = 10))
summary(whole_ranges_location)
```
## output the T6SS ranges data 
```{r}
write.table(whole_ranges_location, file = "../../Results/T6SS_identification/whole_range_of_T6SS.csv", sep = ",", row.names = F)
```


# select the plots that matching the CJIE_determing results

```{r}
CJIE_result <- read.csv("../../Results/CJIE_determining/estimated_CJIE_region.csv")
CJIE_list <- CJIE_result$accession

for (i in CJIE_list) {
  
  file_name <- paste0("../../Results/CJIE_determining/T6SS_with_CJIE_plots/",i,".png")
  
  ggsave(file_name, plot = range_T6SS_plots[[i]],width = 12, height = 4, device = "png")
}
```

