---
title: "CJIE_post_blast_analysis_2"
author: "Hongxi"
date: "`r Sys.Date()`"
output: html_document
---
# Introduction
this document are developed to complete 2 things: 

1) based on the blast result for the CJIE 3 region from RM1221, determine the potential CJIE3 regions in all genomes with assembly level "COMPLETE GENOME".   
2) based on the blast result for three key genes in RM1221 and the regions determined in 1), find out all potential CJIE3 regions that carry any CJIE key genes.

#determine the CJIE3 region from CJIE-3 blast result
##data preparation 
load required packages
```{r message=FALSE, warning=FALSE}
library(tidyr)
library(readxl)
library(gggenes)
library(ggplot2)
library(ggfittext)
library(dplyr)
library(IRanges)
```

##load and wash individual CJIE blast data

```{r}
rm(list = ls())

CJIE_blast_raw <- read.csv("../../Results/CJIE_determining/CJIE_blast_result.csv", sep = " ")
locus_info <- read.csv("../../data/processed_data/CJ_RefSeq/all_locus_info.csv")
CJIE_positive_list <- read.csv("../../Results/summary/complete_genomes_CJIE_positive_list.csv")

CJIE_blast_raw$Subject_accession_version <- sub("\\..*", "", CJIE_blast_raw$Subject_accession_version)
CJIE_blast_raw <- merge(CJIE_blast_raw, locus_info, by.x = "Subject_accession_version", by.y = "locus_accession", all.x = T)

CJIE_positive_in_complete_genome <- merge(CJIE_blast_raw,CJIE_positive_list, by.x = "accession", by.y = "Assembly_Accession")
```

## wash the data to have correct direction for each hits
```{r}

CJIE_positive_in_complete_genome$start_in_genome <- CJIE_positive_in_complete_genome$Start_of_alignment_in_subject + CJIE_positive_in_complete_genome$start
CJIE_positive_in_complete_genome$end_in_genome <- CJIE_positive_in_complete_genome$End_of_alignment_in_subject + CJIE_positive_in_complete_genome$start

CJIE_within_complete_genome <- subset(CJIE_positive_in_complete_genome,
                         subset = Percentage_of_identical_matches >=50,
                         select = c("accession", "Subject_accession_version","start_in_genome","end_in_genome"))

CJIE_within_complete_genome$orientation <- ifelse(CJIE_within_complete_genome$end_in_genome < CJIE_within_complete_genome$start_in_genome, "reverse", "forward")
CJIE_within_complete_genome$type <- "CJIE"

colnames(CJIE_within_complete_genome) <- c("accession","subject_code","start","end","orientation","type")

#build a function to make sure start point < end point
swap_start_end <- function(data) {
  # get the rows that need to be swiped 
  rows_to_swap <- data$orientation == "reverse"
  
  # swap the start and end value 
  data[rows_to_swap, c("start", "end")] <- data[rows_to_swap, c("end", "start")]
  
  # return the swiped data
  return(data)
}

# apply the function
CJIE_within_complete_genome <- swap_start_end(CJIE_within_complete_genome)


CJIE_with_complete_genome_list <- split(CJIE_within_complete_genome,CJIE_positive_in_complete_genome$accession)
```
the blast result with location in subject genomes are storied in "CJIE_with_complete_genome_list" list
NOTE: only 47 genomes are selected from the blast result, which is less than the number (48) of CJIE+ genomes determined by three key genes' blast: indicate that the one contains little coverage of CJIE3 region

## group each hit in genomes into different groups by the criteria of: the gaps of any corrensponding hits larger than 70kb will be considered as different groups   
```{r echo=FALSE}
find_the_ranges <- function(data) {
  all_ranges <- IRanges(start = data$start, end = data$end, names = data$accession)
  ranges <- data.frame(reduce(all_ranges))
  ranges$accession <- unique(data$accession)
  
  threshold <- 70000
  group_the_ranges <- function(df,threshold) {
    accession <- unique(df$accession)
    df <- arrange(df,start)
    df$type <- "CJIE"
    df$group <- cumsum(c(1, (df$start[-1] - df$end[-nrow(df)]) > threshold))
    return(df)
  }
  range_CJIE_list <- group_the_ranges(ranges, threshold)
  
  return(range_CJIE_list)
}

range_CJIE_grouped <- lapply(CJIE_with_complete_genome_list,find_the_ranges)
range_CJIE_grouped <- range_CJIE_grouped[!sapply(range_CJIE_grouped, is.null)]

```
  This function `find_the_ranges` takes a data frame with start, end, and accession columns as input.
  It calculates the ranges and groups them based on a threshold value: 70000 bases.
  the groups numbers more than 2 is rule out from the result
  The function returns a data frame with the grouped ranges.
then apply the function for all elements in list `CJIE_with_complete_genome_list`


### take a view of the possible CJIE3 location (optional)
```{r}
draw_CJIE <- function(df) {
  plot <- ggplot(df, aes(xmin = start, xmax = end,y = accession,label = group)) +
    geom_gene_arrow() +
    geom_gene_label(align = "left") +
    facet_grid(cols = vars(group), scales = "free", space = "fixed") +
    scale_fill_brewer(palette = "Set3",name = "sequence_code") +
    theme_genes()
  return(plot)
}

# apply the function
range_CJIE_plots <- lapply(range_CJIE_grouped,draw_CJIE)

```

### view one of the CJIE range

```{r}
print(range_CJIE_plots["GCF_029025025.1"])
```
##combine all the JICE ranges into a large range and store into one dataframe
```{r}
find_all_ranges <- function(data) {
  
  accession_gene_list <- split(data,data$group)
  
  find_the_ranges <- function(df) {
    seq_ranges <- data.frame(accession = unique(df$accession), start = min(df$start), end = max(df$end), type = unique(df$type) , group = unique(df$group))
    return(seq_ranges)
  }
  
  accession_range_list <- lapply(accession_gene_list, find_the_ranges)
  accession_range_list <- do.call(rbind,accession_range_list)
  
  return(accession_range_list)
}

whole_CJIE_location <- lapply(range_CJIE_grouped,find_all_ranges)
whole_CJIE_location <- do.call(rbind,whole_CJIE_location)
rownames(whole_CJIE_location) <- NULL
whole_CJIE_location <- whole_CJIE_location[c("accession","group","start","end","type")]
```

## output the potential CJIE3 ranges
```{r}
write.table(whole_CJIE_location, file = "../../Results/CJIE_determining/whole_range_of_CJIE.csv", sep = ",", row.names = F)

```



# use CJIE3 key genes' location to determine witch CJIE range to be chosen

## load the CJIE3 key gene blast result data 
```{r}
rm(list = ls())

locus_info <- read.csv("../../data/processed_data/CJ_RefSeq/all_locus_info.csv")
CJIE_positive_list <- read.csv("../../Results/summary/complete_genomes_CJIE_positive_list.csv")
CJIE_gene_blast_raw <- read.csv("../../Results/CJIE_determining/CJIE_key_genes_blast_result.csv", sep = " ")
whole_CJIE_location <- read.csv("../../Results/CJIE_determining/whole_range_of_CJIE.csv")

CJIE_gene_blast_raw$Subject_accession_version <- sub("\\..*", "", CJIE_gene_blast_raw$Subject_accession_version)
CJIE_gene_blast_raw <- merge(CJIE_gene_blast_raw, locus_info, by.x = "Subject_accession_version", by.y = "locus_accession", all.x = T)
CJIE_gene_blast_raw <- merge(CJIE_gene_blast_raw,CJIE_positive_list, by.x = "accession", by.y = "Assembly_Accession")

```
 
## analysis the blast data 
```{r}
# calculate the hits' coverage proportion in query protein 
  CJIE_gene_blast_raw$hit_coverage <- CJIE_gene_blast_raw$Alignment_length / CJIE_gene_blast_raw$Query_sequence_length

# select the qualified results from blast results
  range_CJIE_gene <- subset(CJIE_gene_blast_raw,
                                      subset = Percentage_of_identical_matches >=50 & hit_coverage >= 0.5) 
  
  range_CJIE_gene$start_in_genome <- range_CJIE_gene$Start_of_alignment_in_subject + range_CJIE_gene$start
  range_CJIE_gene$end_in_genome <- range_CJIE_gene$End_of_alignment_in_subject + range_CJIE_gene$start
  
  range_CJIE_gene$orientation <- ifelse(range_CJIE_gene$end_in_genome < range_CJIE_gene$start_in_genome, "reverse", "forward")

#build a function to make sure start point < end point
swap_start_end <- function(data) {
  # get the rows that need to be swiped 
  rows_to_swap <- data$orientation == "reverse"
  
  # swap the start and end value 
  data[rows_to_swap, c("start_in_genome", "end_in_genome")] <- data[rows_to_swap, c("end_in_genome", "start_in_genome")]
  
  # return the swiped data
  return(data)
}

# apply the function
  range_CJIE_gene <- swap_start_end(range_CJIE_gene)
  
  range_CJIE_gene <- subset(range_CJIE_gene, 
                            select = c("accession","Query_accession_version","start_in_genome","end_in_genome"))

  range_CJIE_gene$type <- "key_genes"
  colnames(range_CJIE_gene) <- c("accession","group","start","end","type")
  

```

##conbine the CJIE3 gene location data with potential CJIE region data and try to find the correct CJIE range
```{r}
# combine the two dataframe  
  range_CJIE_all <- rbind(whole_CJIE_location,range_CJIE_gene)
  range_CJIE_all_list <- split(range_CJIE_all,range_CJIE_all$accession)

  find_the_covers <- function(df) {
    accession <- unique(df$accession)
    
    CJIE <- subset(df, type == "CJIE")
    genes <- subset(df, type == "key_genes")
    
    CJIE_range <- IRanges(start = CJIE$start, end = CJIE$end)
    gene_range <- IRanges(start = genes$start, end = genes$end)
    
    cover_status <- CJIE_range %over% gene_range
    
      if (all(cover_status == F)) {
        return(NULL)
      } else {
        ture_CJIE <- CJIE_range[cover_status, ]
        ture_CJIE <- data.frame(ture_CJIE)
        ture_CJIE$accession <- accession 
        ture_CJIE$group <- 1:nrow(ture_CJIE)
        ture_CJIE$type <- "CJIE"
        
        return(ture_CJIE)
      }
    
  }
  
  ture_CJIE_ranges <- lapply(range_CJIE_all_list, find_the_covers)
  ture_CJIE_ranges <- ture_CJIE_ranges[!sapply(ture_CJIE_ranges, is.null)]
  
  CJIE_ranges <- do.call(rbind,ture_CJIE_ranges)
  CJIE_ranges <- CJIE_ranges[c("accession","start","end","width","group","type")]
  row.names(CJIE_ranges) <- NULL
```

##output the confirmed CJIE range information
```{r}
write.csv(CJIE_ranges, file = "../../Results/CJIE_determining/confirmed_CJIE_ranges.csv",row.names = F)
```


