---
title: "select_CJIE_covered_locus.Rmd"
author: "Hongxi"
date: "`r Sys.Date()`"
output: html_document
---

this code is used to select determined CJIE regions from respective genomes files. 
input files are 
1. Determined CJIE ranges, with the information of:
    1) accession 
    2) start point in genome
    3) end point in genome
    3) group NO. for each CJIE in their genome (e.g. group = 2 means this regions is the second potenital CJIE3 region in this accession)
2. locus information collected from each .gbk file

# determine the location should be selected on each locus for CJIE-3 region
##load packages
```{r message=FALSE, warning=FALSE}
library(Biostrings)
library(dplyr)
```
##load the data 
```{r}
rm(list = ls())

CJIE_ranges <- read.csv("../../Results/CJIE_determining/potential_CJIE_ranges.csv")
locus_info <- read.csv("../../data/processed_data/CJ_RefSeq/all_locus_info.csv")

CJIE_list <- CJIE_ranges["accession"]

locus_info_CJIE <- merge(locus_info,CJIE_list,by = "accession", all.y =T)

locus_info_CJIE$type <- "locus"
```

## wash the data for next step
```{r}
CJIE_ranges <- CJIE_ranges[c("accession","group","width","start","end","type")]
colnames(CJIE_ranges) <- c("accession","locus_accession","length","start","end","type")

locus_and_CJIE <- rbind(locus_info_CJIE,CJIE_ranges) 

locus_and_CJIE_list <- split(locus_and_CJIE,locus_and_CJIE$accession)

```

##write a function to determine what position should be select in each locus

```{r}
find_locus_location <- function(data) {
  
  locus <- subset(data, type == "locus")
  CJIE <- subset(data, type == "CJIE")
  
  mutiple_CJIE <- split(CJIE,CJIE$locus_accession)
  
  find_location <- function(df,locus) {
      group_num <- df$locus_accession
      target_start <- df$start
      target_end <- df$end
      
      
      locus$target_start_in_locus <- ifelse(locus$start >= target_start, 1, target_start-locus$start)
      locus$target_end_in_locus <- ifelse(locus$end >= target_end, target_end-locus$start, locus$length)
      locus <- subset(locus,target_start_in_locus <= length & target_end_in_locus > 0)
      
      locus$group <- group_num
      return(locus)
  }
    location_on_locus <- lapply(mutiple_CJIE, find_location, locus = locus)
    location_on_locus <- location_on_locus[!sapply(location_on_locus, is.null)]
    
    return(location_on_locus)
}

location_on_locus <- lapply(locus_and_CJIE_list, find_locus_location)

location_on_locus_final <- unlist(location_on_locus, recursive = F)
location_on_locus_final <- do.call(rbind, location_on_locus_final)

#wash the data
location_on_locus_final <- distinct(location_on_locus_final)
rownames(location_on_locus_final) <- NULL
```
now the position for each locus that needed to be cut are stored in "location_on_locus_final" 

#cat the target sequence from downloaded files
##load the packages
```{r message=FALSE}
library(Biostrings)
```

##load and wash the files
```{r}
folder_path <- "../../data/processed_data/CJ_RefSeq/ncbi_dataset/data/"

genome_and_locus_info <- location_on_locus_final
genome_and_locus_info$target_start_in_locus <- as.integer(genome_and_locus_info$target_start_in_locus)

```

##get the target sequences
```{r}


get_the_file_path <- function(accession,data_folder_path) {
  genome_folder_path <- paste0(data_folder_path,accession)
  genome_path <- list.files(path = genome_folder_path, pattern = "^GCF_.*\\.fna$", full.names = T)
  return(genome_path)
}

genome_and_locus_info$genome_file_path <- get_the_file_path(accession = genome_and_locus_info$accession, data_folder_path = folder_path)

genome_and_locus_info$accession_group <- paste0(genome_and_locus_info$accession,"_",genome_and_locus_info$group)

genome_and_locus_list <- split(genome_and_locus_info, genome_and_locus_info$accession_group)


load_and_select_genome <- function(df) {
  accession <- unique(df$accession_group)
  genome_path <- unique(df$genome_file_path)
  
  genome_seq <- readDNAStringSet(genome_path)
  
  #rename the locus to fit locus name in dataframe
  seq_names <- names(genome_seq)
  new_seq_names <- sapply(seq_names, function(x) unlist(strsplit(x, "\\."))[1])
  names(genome_seq) <- new_seq_names
  
  df <- df[order(df$start),]
  df$sequence <- subseq(genome_seq[df$locus_accession], start = df$target_start_in_locus, end = df$target_end_in_locus)
  
  sequence_final <- paste0(df$sequence,collapse = "")
  
  start_in_genome <- min(df$target_start_in_locus + df$start)
  end_in_genome <- max(df$target_end_in_locus + df$start)
  
  data <- data.frame(accession = accession, 
                     CJIE_seq = sequence_final,
                     start = start_in_genome,
                     end = end_in_genome)
  
  return(data)
  
}

CJIE_regions_final <- lapply(genome_and_locus_list,load_and_select_genome)
CJIE_regions_final <- CJIE_regions_final[!sapply(CJIE_regions_final, is.null)]
CJIE_regions_final_table <- do.call(rbind, CJIE_regions_final)

rownames(CJIE_regions_final_table) <- NULL
```
the sequence for each CJIE regions are storied in the list `CJIE_regions_final`

##analysis each CJIE-3 regions
```{r}
#get the GC content
CJIE_regions_final_table$GC_content <- sapply(CJIE_regions_final_table$CJIE_seq, function(x) {
    sum(letterFrequency(DNAString(x), c("G", "C"))) / length(DNAString(x))
})

#size for each CJIE-3 region
CJIE_regions_final_table$size <- CJIE_regions_final_table$end -CJIE_regions_final_table$start

#wash the table 
CJIE_regions_final_table <- CJIE_regions_final_table[c("accession","start","end","size","GC_content")]
```

##output the CJIE-3 regions
```{r}
write.csv(CJIE_regions_final_table, file = "../../Results/CJIE_determining/potential_CJIE_ranges_with_sequence.csv", row.names = F)

outputdir <- "../../data/processed_data/CJIE-3_regions/"

outputCJIE <- function(df,direction) {
  accession <- df$accession
  
  textout <- paste0(">",accession,"\n",df$CJIE_seq)
  
  outpath <- paste0(direction,accession)
  dir.create(outpath,showWarnings = FALSE)
  
  outname <- paste0(outpath,"/CJIE-3.fna")
  writeLines(textout, outname)
  
}

lapply(CJIE_regions_final, outputCJIE, direction = outputdir)
```
then apply the prokka to annotate the CJIE-3 region

##test code, not available
```{r eval=FALSE, include=FALSE}

testdata <- genome_and_locus_list[["GCF_012220765.1_1"]]

accession <- unique(testdata$accession)
  genome_path <- unique(testdata$genome_file_path)
  
  genome_seq <- readDNAStringSet(genome_path)
  
  #rename the locus to fit locus name in dataframe
  seq_names <- names(genome_seq)
  new_seq_names <- sapply(seq_names, function(x) unlist(strsplit(x, "\\."))[1])
  names(genome_seq) <- new_seq_names
  
  testdata <- testdata[order(testdata$start),]
  testdata$sequence <- subseq(genome_seq[testdata$locus_accession], start = testdata$target_start_in_locus, end = testdata$target_end_in_locus)

  sequence_final <- paste0(testdata$sequence, collapse = "")

  testdata2 <- data.frame(accession = accession, 
                     CJIE_seq = sequence_final)
  
  
  genome_seq
test_Sequence <- xscat(test_genome$sequence)

substr(sequence_final, nchar(sequence_final) - 99, nchar(sequence_final))


```