---
title: "generate_the_graph.Rmd"
author: "Hongxi"
date: "`r Sys.Date()`"
output: html_document
---

this file is used to generate the genetic structure of CJIE-3 region from the annotated files

##load the library
```{r warning=FALSE}
library(dplyr)
library(stringr)
library(ggplot2)
library(gggenes)
library(ggrepel)
```

##load the data
```{r}
rm(list = ls())

type_group <- read.csv("./protein_function_and_type.csv") 
CJIE_region_info <- read.csv("../../Results/CJIE_determining/potential_CJIE_ranges.csv")
CJIE_region_info$accession <- paste0(CJIE_region_info$accession,"_",CJIE_region_info$group)


```

##list the file path
```{r}
data_folder <- "../../data/processed_data/CJIE-3_regions"
folders_link <- list.files(data_folder, pattern = "\\.gff$", recursive = T, full.names = T)
folders_link <- as.list(folders_link)

# name each link in the list
folder_names <- character(length(folders_link))
for (i in seq_along(folders_link)) {
    folder_names[[i]] <- basename(dirname(dirname(folders_link[[i]])))
}
names(folders_link) <- folder_names
```


#make a function to extract cds information from annotated .gff files
```{r}

extract_cds <- function(file_link, CJIE_info) {
  #wash gff file
  data <- readLines(file_link)
  
  comment_lines <- grep("^##",data)
  tables <- data[(comment_lines[2]+1):(comment_lines[3]-1)]
  
  df <- str_split_fixed(tables, "\t",9)
  df <- data.frame(df)
  colnames(df) <- c("accession","name","tag","start_in_locus","end_in_locus","V6","orientation", "V8", "protein_function")
  df <- df[c("accession","tag","start_in_locus","end_in_locus","orientation","protein_function")]
  df$start_in_locus <- as.numeric(df$start_in_locus)
  df$end_in_locus <- as.numeric(df$end_in_locus)
  df$protein_function <- gsub(".*product=", "", df$protein_function)
  
  df <- left_join(df,type_group,by = "protein_function")
  df$type <- ifelse(is.na(df$type),"other",df$type)
  df$color <- ifelse(is.na(df$color), 'yellow', df$color)
  
  #take CJIE information
  accession_NUM <- unique(df$accession)
  
  CJIE <- subset(CJIE_info,subset = accession == accession_NUM, select = c("accession","start","end"))
  start_in_genome <- CJIE$start
  end_in_genome <- CJIE$end
  
  #add the CJIE location information to df
  df$start <- df$start_in_locus + start_in_genome
  df$end <- df$end_in_locus + start_in_genome
  
  return(df)
} 

cds_info <- lapply(folders_link, extract_cds, CJIE_info = CJIE_region_info)
```
now all the CDS information are stored in cds_info


##generate the graphs
```{r}
draw_CJIE <- function(df) {
  feature <- subset(df,
                       subset = protein_function != "hypothetical protein" & type != "T6SS_core",
                       select = c("accession","orientation","protein_function","type","start","end"))
  
  CJIE_plot <- ggplot(df,
                      aes(xmin = start, 
                          xmax = end, 
                          y = orientation, 
                          fill = type, 
                          label = type, 
                          forward = orientation == "+")) +
    geom_gene_arrow() +
    geom_text_repel(data = feature,
                    aes(x = (start+end)/2, 
                        y = orientation, 
                        label = protein_function),
                    direction = "y",
                    force_pull = 0.2,
                    nudge_y = 0.3,
                    max.overlaps = 20,
                    segment.linetype = "dashed",
                    segment.size = 0.1,
                    size = 3) +
    facet_grid(scales = "free", space = "fixed") +
    scale_fill_manual(values = setNames(df$color,df$type)) +
    theme_genes()+
    xlab("")+
    ylab(unique(df$accession))
  return(CJIE_plot)
}

all_CJIE_plots <- lapply(cds_info,draw_CJIE)
```


###have a look of one of the plot(optional)
```{r}
print(all_CJIE_plots[[1]])
```

##output the plots into respective folder
```{r}
out_folder <- "../../Results/CJIE_determining/CJIE3_plots/"

for (i in seq_along(all_CJIE_plots)) {
  out_path <- paste0(out_folder,names(all_CJIE_plots[i]))
  ggsave(out_path,
         plot = all_CJIE_plots[[i]],
         width = 20,
         height = 5,
         device = "png")
}
```

###test code
```{r eval=FALSE, include=FALSE}

data <- readLines("/home/vader/Documents/MSc-project/data/processed_data/CJIE-3_regions/GCF_001717625.1_1/annotation/PROKKA_08182023.gff")

comment_lines <- grep("^##", data)

result <- data[(comment_lines[2]+1):(comment_lines[3]-1)]

df <- str_split_fixed(result, "\t",9)
df <- data.frame(df)
colnames(df) <- c("accession","name","tag","start_in_locus","end_in_locus","V6","orientation", "V8", "protein_function")

df$protein_function <- gsub(".*product=", "", df$protein_function)

df <- df[c("accession","tag","start_in_locus","end_in_locus","orientation","protein_function")]
df$start_in_locus <- as.numeric(df$start_in_locus)
df$end_in_locus <- as.numeric(df$end_in_locus)

testdata <- left_join(df,type_group,by = "protein_function")

testdata$type <- ifelse(is.na(testdata$type),"other",testdata$type)
testdata$color <- ifelse(is.na(testdata$color), 'yellow', testdata$color)


accession_NUM <- unique(testdata$accession)

CJIE <- subset(CJIE_region_info,subset = accession == accession_NUM, select = c("accession","start","end"))
start_in_genome <- CJIE$start
end_in_genome <- CJIE$end

testdata$start <- testdata$start_in_locus + start_in_genome
testdata$end <- testdata$end_in_locus + start_in_genome

test_feature <- subset(testdata,
                       protein_function != "hypothetical protein" & type != "T6SS_core",
                       select = c("accession","orientation","protein_function","type","start","end"))



CJIE_plot <- ggplot(testdata,aes(xmin = start, xmax = end, y = orientation, fill = type, label = type, forward = orientation == "+")) +
  geom_gene_arrow() +
  geom_text_repel(data = test_feature,aes(x = (start+end)/2, y = orientation, label = protein_function),
                  direction = "y",
                  force_pull = 0.2,
                  nudge_y = 0.3,
                  max.overlaps = 20,
                  segment.linetype = "dashed",
                  segment.size = 0.1,
                  size = 3) +
  facet_grid(scales = "free", space = "fixed") +
  scale_fill_manual(values = setNames(testdata$color,testdata$type)) +
  theme_genes()+
  xlab("")+
  ylab(unique(testdata$accession))
CJIE_plot
```

